Kaggle-in-class Data Challenges Can Boost Student Learning

### 2019-09-23 paper-kaggle.Rmd
teachers experience

## Working with the Kaggle InClass platform

Creating a new competition is surprisingly easy. 
Information on setting up a Kaggle InClass challenge can be found at https://www.kaggle.com/about/inclass/overview. 
To ease the setting process, we created a short video illustrating the steps to establish a competition....

Finding data for a competition may take some time. It is not recommended to take freely available data sets from the Internet. A clever student will always try to check if the data is on the web. If it is, she will use the answers from the data set instead of predicting the answers with a self-developed model. As a statistician, and an instructor in a subject, you or one of your colleagues usually have some interesting data set to use. In such a case, the data set and particularly the target values will be unfamiliar to the students.

Once you open the competition, it runs without any intervention from the instructor. The instructor can monitor students progress: the number of submissions, student scores and even the uploaded data at any time. When the competition ends the "Leaderboard" page provides a list of students ordered by the final score. It also provides all the scores from all past submissions (under "Raw Data" on "Public Leaderboard").

From our experience, we recommend limiting the number of daily submissions to one. It encourages students to think about more efficient improvement of their model before the next submissions.  We also believe that it is better to start the competition in the second half of the semester and finish it two to one week before the end of the semester. It allows to cover many different statistical methods that students may apply during the competition. Finishing some time before the end of the semester reduces students anxiety at the end of the semester and the approaching exams.

@@

>>>>

@@

# Teachers' corner

Creating a new competition is surprisingly easy. Information on setting up a Kaggle InClass challenge is available on the service's web site (https://www.kaggle.com/about/inclass/overview). There is a setup wizard for step-by-step guidance on getting your competition underway. We have created a short video illustrating the steps to establish a new competition, and this is provided in supplementary material for the paper.

Finding a suitable data set for a competition can be a challenge. The criteria for a good data set are:

1. the full set is not available to the students, to avoid cheating.
2. the data is not too easy, or too hard, to model, so that there is some discriminatory ower in the results.
3. contains some challenges, that make standard off-the-shelf modeling less successful, like different variable types that need processing or transforming, some outliers, large number of variables. 
4. If it is a classification challenge, it will work better with relatively balanced classes, because the overall accuracy is the easiest metric to use.

<!--It is not recommended to take freely available data sets from the Internet. A clever student will always try to check if the data is on the web. If it is, she will use the answers from the data set instead of predicting the answers with a self-developed model. As a statistician, and an instructor in a subject, you or one of your colleagues usually have some interesting data set to use. In such a case, the data set and particularly the target values will be unfamiliar to the students.
-->

Choosing the metric upon which to evaluate the model is another decision. Our advice is to keep it simple, so you, and the students, can understand the student scores. If you are running a regression challenge, then "Root Mean Squared Error (RMSE)". It is is a classification challenge, then "Categorization Accuracy", the percentage correct is reasonable.

The data needs to be split into training and testing sets. Kaggle will then split your test set into two, a public set that is used to provide ongoing scores to participants, and a private set, on which performance is revealed only after the competition closes. If you have categorical variables in the data set, you will want to make sure that all categories are present in both training and test sets. The training set will have both predictors and response, but the test set will have the response variable removed. Each observation needs to be assigned an id, because this will be needed to evaluate predictions. The solution file, containing the id and the true response, is provided to the system for evaluating submissions, and is kept private. A sample submission file needs to be provided. Participants will submit their solutions in the same format.

It is a good idea to build a basic model yourself on the training data, and predict the test data. The performance of this model can be provided to the participants as baseline to beat.

The competition needs to run without any intervention from the instructor. The instructor can monitor students progress: the number of submissions, student scores and even the uploaded data at any time. When the competition ends the "Leaderboard" page provides a list of students ordered by the final score. It also provides all the scores from all past submissions (under "Raw Data" on "Public Leaderboard").

It may be recommended to limit students to one submission per day. It encourages students to think about more efficient improvement of their model before the next submission. It also prevents the student spending too much time building and submitting models. Some students will become so engaged in the competition tha they might neglect their other coursework. About halfway throug hthe coompetition 
<!--We also believe that it is better to start the competition in the second half of the semester and finish it two to one week before the end of the semester. It allows to cover many different statistical methods that students may apply during the competition. Finishing some time before the end of the semester reduces students anxiety at the end of the semester and the approaching exams. -->

In awarding course points to student effort, we typically align it to performance. Participant ranks based on their performance on the private part of the test data are recorded. Performance scores that are pretty close to each other should be given the same rank, reflecting that there may not be a significant difference between them. The best gets perhaps, 5 points, then a half a point drop until about 2.5 points, so that the worst performing students still get 50% for the task. Kaggle does not allow you to download participants email addresses. All you see is their kaggle name. Record the student names in kaggle to match with your class records.

From an instructor perspective, its very rewarding watching the students participate in the competition. It provides a truly objective way to assess their ability to model in practice. Students are often motivated to consult with the instructor about why their model is underperforming, or what other approaches might produce better results.
